Automatic conversion of colloquial Finnish to standard Finnish

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
add more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58603 72bbbca6-d526-0410-a7d9-f06f51895060

\begin{itemize}
  \item 273 rules from \newcite{karlsson2008}'s grammar of Finnish (\S95--97). The 
    rules took around one hour to implement.
  \item 98 rules written by examining the development corpus, these rules also
    took approximately one hour to implement.
\end{itemize}

@@

\begin{table*}
  \centering
  \begin{tabular}{|l|r|l|l|}
    \hline
    \textbf{Input:} & \multicolumn{3}{l|}{M\"{a} oon Tomminkaa `I am with Tommi'.} \\
    \hline
    \multirow{2}{*}{Step 1} & & M\"{a} oon Tomminkaa & apply rule 1 \\
                            & & Min\"{a} oon Tomminkaa & m\"{a} $\rightarrow$ min\"{a} \\
    \hline
    \multirow{2}{*}{Step 2} & & M\"{a} oon Tomminkaa & apply rule 2 \\
                            &   & Min\"{a} oon Tomminkaa & oon $\rightarrow$ olen \\
                            &   & M\"{a} olen Tomminkaa & \\
                            &   & Min\"{a} olen Tomminkaa & \\
    \hline
    \multirow{2}{*}{Step 3} &   & M\"{a} oon Tomminkaa   & apply rule 3\\
                            &   & Min\"{a} oon Tomminkaa  & (?+)nkaa $\rightarrow$ \textbackslash1n kanssa \\
                            &   & M\"{a} olen Tomminkaa  &\\
                            &   & Min\"{a} olen Tomminkaa  &\\
                            &   & M\"{a} oon Tommin kanssa  &\\
                            &   & Min\"{a} oon Tommin kanssa  &\\
                            &   & M\"{a} olen Tommin kanssa  &\\
                            &   & Min\"{a} olen Tommin kanssa  &\\
    \hline
    \multirow{2}{*}{Step 4} & -4.5811 &  Min\"{a} olen Tommin kanssa & rank candidates \\
                            & -7.8174 &  Min\"{a} oon Tommin kanssa & \\
                            & -8.0941 &  M\"{a} olen Tommin kanssa & \\
                            & -8.8651&  M\"{a} oon Tommin kanssa & \\
                            & -9.2045 &  Min\"{a} olen Tomminkaa & \\
                            & -12.4408 &  Min\"{a} oon Tomminkaa & \\
                            & -12.7176 &  M\"{a} olen Tomminkaa & \\
                            & -13.4885 &  M\"{a} oon Tomminkaa & \\

@@

We trained models based on two approaches, the first being phrase-based machine translation
\cite{zens2002} and the second on character-based machine translation \cite{nakov2012,TiedemannEAMT2009}.

@@

In order to evaluate how well the different normalisation strategies worked in 
combination with another language technology tool, we performed an end-to-end experiment
involving machine translation.
To evaluate this, we took the test corpus and translated it to English.

We then translated the corpus with each of the best performing methods of each class, and 
calculated the {\sc bleu} score.

@@

\begin{table}
  \centering
  \begin{tabular}{|l|r|r|r|}
     \hline
    \textbf{System} & \textbf{PER} & \textbf{WER} & \textbf{BLEU} \\
     \hline
     Baseline & 46.12 & 48.04 & 26.31 \\
     \hline
     rules-1 & 38.27 & 41.19 & 32.65 \\
     rules-2 & 38.17 & 35.25 & 36.41 \\
     rules-c & 36.56 & 39.68 & 34.68 \\
     \hline
     CBMT-cn & 43.09 & 48.34 & 33.55 \\
     CBMT-cs & 46.22 & 52.27 & 29.21 \\
     PBMT-cn & 28.05 & 35.42 & 48.37 \\
     PBMT-cs & 27.95 & 36.13 & 46.76 \\
     \hline
  \end{tabular}
  \caption{Results for the normalisation task}
  \label{table:results-norm}
\end{table}

@@

\begin{table}
  \centering
  \begin{tabular}{|l|r|r|r|}
     \hline
    \textbf{System} & \textbf{PER} & \textbf{WER} & \textbf{BLEU} \\
     \hline
     Colloquial & - & - & 11.95 \\
     Normalised & - & - & 21.81 \\
     Standardised & - & - & 17.62 \\
     \hline
     rules-2 & - & - & 12.28 \\
     cbmt-cn & - & - & 12.35 \\
     pbmt-cn & - & - & 17.90 \\
     \hline
  \end{tabular}
  \caption{Results for the translation task}
  \label{table:results-trad}
\end{table}

>>>>

\begin{itemize}
  \item \texttt{rules-1}: 273 rules from \newcite{karlsson2008}'s grammar of Finnish (\S95--97). The 
    rules took around one hour to implement.
  \item \texttt{rules-2}: 98 rules written by examining the development corpus, these rules also
    took approximately one hour to implement.
\end{itemize}

@@

\begin{table*}
  \centering
  \begin{tabular}{|l|r|l|l|}
    \hline
    \textbf{Input:} & \multicolumn{3}{l|}{M\"{a} oon Tomminkaa `I am with Tommi'.} \\
    \hline
    \multirow{2}{*}{Step 1} & & M\"{a} oon Tomminkaa & apply rule 1: \\
                            & & Min\"{a} oon Tomminkaa & m\"{a} $\rightarrow$ min\"{a} \\
    \hline
    \multirow{2}{*}{Step 2} & & M\"{a} oon Tomminkaa & apply rule 2: \\
                            &   & Min\"{a} oon Tomminkaa & oon $\rightarrow$ olen \\
                            &   & M\"{a} olen Tomminkaa & \\
                            &   & Min\"{a} olen Tomminkaa & \\
    \hline
    \multirow{2}{*}{Step 3} &   & M\"{a} oon Tomminkaa   & apply rule 3:\\
                            &   & Min\"{a} oon Tomminkaa  & (?+)nkaa $\rightarrow$ \textbackslash1n kanssa \\
                            &   & M\"{a} olen Tomminkaa  &\\
                            &   & Min\"{a} olen Tomminkaa  &\\
                            &   & M\"{a} oon Tommin kanssa  &\\
                            &   & Min\"{a} oon Tommin kanssa  &\\
                            &   & M\"{a} olen Tommin kanssa  &\\
                            &   & Min\"{a} olen Tommin kanssa  &\\
    \hline
    \multirow{2}{*}{Step 4} & -4.5811 &  Min\"{a} olen Tommin kanssa & rank candidates \\
                            & -7.8174 &  Min\"{a} oon Tommin kanssa & \\
                            & -8.0941 &  M\"{a} olen Tommin kanssa & \\
                            & -8.8651&  M\"{a} oon Tommin kanssa & \\
                            & -9.2045 &  Min\"{a} olen Tomminkaa & \\
                            & -12.4408 &  Min\"{a} oon Tomminkaa & \\
                            & -12.7176 &  M\"{a} olen Tomminkaa & \\
                            & -13.4885 &  M\"{a} oon Tomminkaa & \\

@@

We trained models based on two approaches, the first being phrase-based machine translation (PBMT, \newcite{zens2002}) and the second on character-based machine translation (CBMT, \newcite{nakov2012,TiedemannEAMT2009}).

@@

In order to evaluate how well the different normalisation strategies worked in 
combination with another language technology tool, we performed an end-to-end experiment
involving machine translation.
To evaluate this, we took the colloquial portion of the test corpus and manually 
translated it to English. For each of the best-performing systems we first passed
the colloquial text through, and then translated the output to English using
a widely-used online machine translation engine with Finnish to English. We compared
the output to translating the text to English without the standardiser.

@@

\begin{table}
  \centering
  \begin{tabular}{|l|r|r|r|}
     \hline
    \textbf{System} & \textbf{PER} & \textbf{WER} & \textbf{BLEU} \\
     \hline
     Baseline & 46.12 & 48.04 & 26.31 \\
     \hline
     rules-1 & 38.27 & 41.19 & 32.65 \\
     rules-2 & \textbf{38.17} & \textbf{35.25} & \textbf{36.41} \\
     rules-c & 36.56 & 39.68 & 34.68 \\
     \hline
     CBMT-cn & 43.09 & 48.34 & 33.55 \\
     CBMT-cs & 46.22 & 52.27 & 29.21 \\
     PBMT-cn & \textbf{28.05} & \textbf{35.42} & \textbf{48.37} \\
     PBMT-cs & 27.95 & 36.13 & 46.76 \\
     \hline
  \end{tabular}
  \caption{Results for the normalisation task. The system \texttt{rules-c} is the combination
     of the rules in \texttt{rules-1} and \texttt{rules-2}. The figures in bold are the 
     best results for unsupervised and supervised methods.}
  \label{table:results-norm}
\end{table}

@@

% unsupervised does as well as CBMT, and and half the way towards pbmt

% supervised systems trained on the normalised corpus outperform those trained on the standardised corpus (WHY?)

% the rules written from the development portion of the corpus outperform the rules from the grammar

\begin{table}
  \centering
  \begin{tabular}{|l|r|r|r|}
     \hline
    \textbf{System} & \textbf{PER} & \textbf{WER} & \textbf{BLEU} \\
     \hline
     Colloquial   & 45.31 & 71.28 & 11.95 \\
     Normalised   & 34.63 & 62.34 & 21.81 \\
     Standardised & 38.67 & 63.35 & 17.62 \\
     \hline
     rules-2      & 42.71 & 70.42 & 12.28 \\
     \hline
     cbmt-cn      & 44.44 & 74.17 & 12.35 \\
     pbmt-cn      & 38.38 & 66.81 & 17.90 \\
     \hline
  \end{tabular}
  \caption{Results for the translation task. The first three rows are the results 
    from translating the sections of the corpus. As with the }
  \label{table:results-trad}
\end{table}

% example translations

%colloq: ooh what onks there snow
%pbmt: ooh what if there's snow
%ref: ooh what, is there snow?
